\input{../preamble}

%----------------------------------------------------------------------------------------
% TITLE PAGE
%----------------------------------------------------------------------------------------

\title[第17讲]{第十七讲 ：文件系统} % The short title appears at the bottom of every slide, the full title is only on the title page
\subtitle{第5节：进程文件系统procfs}
\author{向勇、陈渝、李国良} % Your name
\institute[清华大学] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
  清华大学计算机系 \\ % Your institution for the title page
  \medskip
  \textit{xyong,yuchen,liguoliang@tsinghua.edu.cn} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}
\section{第5节：进程文件系统procfs}
% ### 17.5 procfs
% 
%------------------------------------------------
% \begin{frame}
% \frametitle{提纲} % Table of contents slide, comment this block out to remove it
% \tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
% \end{frame}
%------------------------------------------------
\begin{frame}[fragile]
    \frametitle{What is \href{https://en.wikipedia.org/wiki/Procfs}{procfs}?}


    \begin{itemize}
        \item \href{http://dtrace.org/blogs/eschrock/2004/06/25/a-brief-history-of-proc/}{Tom Killian} wrote the first implementation of /proc, explained in his paper published in 1984.
	    \begin{itemize}
	        \item It was designed to replace the venerable ptrace system call
      \end{itemize} \pause
        \item It presents information about processes and other \textcolor{blue}{system information}
	    \begin{itemize}
	        \item Provide a more convenient and standardized method for \textcolor{blue}{dynamically accessing} process data held in the kernel
          \item The proc file system acts as an interface to \textcolor{blue}{internal data structures} in the kernel.
          \item The proc filesystem provides a \textcolor{blue}{method of communication} between kernel space and user space.
      \end{itemize} \pause
        \item The proc filesystem (procfs) is a special filesystem
	    \begin{itemize}
	        \item It is mapped to a mount point named /proc at boot time.
          \item All of them have a file size of 0, with the exception of kcore, mtrr  and self.
          \item it as a window into the kernel. it just acts as a pointer to where the actual  process information resides.
      \end{itemize}
    \end{itemize}
% 
\end{frame}
%------------------------------------------------
\begin{frame}[fragile]
    \frametitle{The purpose and contents of procfs}


    \begin{itemize}
        \item /proc/PID/fd
	    \begin{itemize}
	        \item Directory,  which contains all \textcolor{blue}{file descriptors}.
      \end{itemize}
        \item /proc/PID/maps
	    \begin{itemize}
	        \item \textcolor{blue}{Memory  maps} to executables and library files.
      \end{itemize}
        \item /proc/PID/status
	    \begin{itemize}
	        \item \textcolor{blue}{Process  status} in human readable form.
      \end{itemize}
        \item /proc/mtrr
	    \begin{itemize}
	        \item The \href{https://wiki.osdev.org/MTRR}{Memory Type Range Registers} (MTRRs) may be used to control processor  access to memory ranges.
          \item This is most useful when you have a video (VGA)  card on a PCI or AGP bus.
          \item This can increase performance of image write operations 2.5  times or more.
      \end{itemize}
    \end{itemize}
% 
\end{frame}
%------------------------------------------------
\begin{frame}[fragile]
    \frametitle{\href{https://developer.ibm.com/technologies/linux/articles/l-proc/}{Access the Linux kernel using the /proc filesystem}}


    \begin{itemize}
        \item kernel modules
	    \begin{itemize}
	        \item dynamically add or remove code from the Linux kernel.
          \item Inserting, checking, and removing an LKM
          \item Reviewing the kernel output from the LKM
      \end{itemize} \pause
        \item Integrating into the /proc filesystem
	    \begin{itemize}
	        \item To create a virtual file in the /proc filesystem, use the `create\_proc\_entry` function.
          \item To remove a file from /proc, use the `remove\_proc\_entry` function.
          \item Write to a /proc entry (from the user to the kernel) by using a `write\_proc` function.
          \item Read data from a /proc entry (from the kernel to the user) by using the `read\_proc`function.
      \end{itemize}
    \end{itemize}
% 
\end{frame}
%------------------------------------------------
\begin{frame}[fragile]
    \frametitle{Interactive tour of /proc: `ls /proc`}

% 
Reading from and writing to /proc (configuring the kernel): 
% 
\begin{semiverbatim}
cat /proc/sys/net/ipv4/ip\_forward
0
echo "1" > /proc/sys/net/ipv4/ip\_forward
cat /proc/sys/net/ipv4/ip\_forward
1
\end{semiverbatim}
% 
\end{frame}
%------------------------------------------------
\begin{frame}[fragile]
    \frametitle{\href{https://www.cnblogs.com/qiuheng/p/5761877.html}{procfs, sysfs, debugfs in Linux}}


    \begin{itemize}
        \item procfs 历史最早，最初就是用来跟内核交互的唯一方式，用来获取处理器、内存、设备驱动、进程等各种信息。
        \item sysfs 跟 kobject 框架紧密联系，而 kobject 是为设备驱动模型而存在的，所以 sysfs 是为设备驱动服务的。
        \item debugfs 从名字来看就是为debug而生，所以更加灵活。
    \end{itemize}
% 
\end{frame}
%----------------------------------------------
%----------------------------------------------
\end{document}
